---
title: "Landownership Appears Neither Necessary nor Sufficient for Superior Agricultural Performance in Ghana"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from land_tenure.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: Drafting (v001)](https://img.shields.io/badge/status-drafting-lightgrey)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: Drafting (v001 produced 20 May 2026); GAAE 2025 conference paper available.

### Contributors

- [Gilbert Addae](https://github.com/Gilbertaddae)
- [Jacob Asravor](https://scholar.google.com/citations?user=_zUi3FsAAAAJ&hl=en)
- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)

### Abstract

Secure land tenure is widely assumed to stimulate investment and raise agricultural productivity, yet rigorous evidence from sub-Saharan Africa remains inconclusive. Leveraging six rounds of the Ghana Living Standards Survey (GLSS 3-7), this study links land-ownership status to production outcomes for 35,185 farm households. We use statistical matching within a meta-stochastic frontier framework to isolate ownership effects on input elasticities, returns to scale, technology gaps, and technical efficiency. Across the matched sample, all inputs significantly boost output, but fertilizer and pesticide use generate larger marginal gains for non-landowners. Landholders register slightly higher returns to scale (Delta = 0.0154); however, at the aggregate level we detect no statistically significant difference in either technology adoption or technical efficiency between owners and non-owners. Disaggregating by tenure form, documented owners and farmers with transferable or informal rights show lower meta-frontier access and reduced technical efficiency relative to matched non-owners, challenging the conventional narrative that tenure security automatically delivers productivity gains.

**Keywords**: land tenure; ownership; meta-stochastic frontier; technical efficiency; technology gap; Ghana

**JEL Classification**: Q15, Q12, O13, K11

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
