---
output:
  html_document: default
  pdf_document: default
---

```{r setup-00-abstract, include=FALSE}
# Knitting this section on its own? Load what ../land-tenure.Rmd normally
# provides. A no-op under the parent, 303 or 304, which already have it.
if (!exists("fmt_num")) {
  .s <- c("_setup.R", "sections/_setup.R", "narrative/sections/_setup.R")
  .s <- .s[file.exists(.s)]
  if (!length(.s)) stop("cannot find sections/_setup.R from ", getwd(), call. = FALSE)
  source(.s[1])
}
```

# Abstract

Secure land tenure is widely expected to raise agricultural productivity by strengthening investment incentives and improving access to credit, making land titling a common policy across sub-Saharan Africa. However, whether land ownership itself improves farm performance remains unsettled, as aggregate comparisons may conceal differences in the technology farmers can access and the efficiency with which they use it. We address this issue by examining productivity differences between landowners and non-owners in Ghana, with a focus on technology endowment and production efficiency. Combining statistical matching with a meta-stochastic frontier analysis on `r fmt_num(N_ALL, 0)` farm households across five rounds of the Ghana Living Standards Survey, we find negligible differences between landowners and matched non-owners, with gaps of less than one percentage point across all measures, indicating that non-owners have marginally higher access to the best available technology, while landowners use their production frontier more efficiently. Disaggregating ownership further shows that documented owners and farmers with transferable land rights have lower meta-frontier efficiency than comparable non-owners, with the gap widening as transfer rights become more complete. Farmers who acquired land through purchase, rental, or family allocation are closer to the available technology but are less efficient in using it. These findings suggest that land ownership is neither necessary for accessing available technology nor sufficient for using it efficiently, as rental, sharecropping, and borrowing can already facilitate access to land while documentation and transfer rights provide no clear efficiency advantage. Landownership  alone is therefore unlikely to raise agricultural productivity without complementary improvements in input use, credit access, and market conditions.

```{r frontmatter-extras, echo=FALSE, results='asis'}
# Keywords and JEL codes belong in the manuscript at economics journals and are
# supplied through the submission system at others; Nature Communications uses
# neither. Switch with ARTICLE_TARGET, the same idiom as ARTICLE_CSL in the
# parent Rmd: ARTICLE_TARGET=nature drops both lines, anything else keeps them.
if (!identical(tolower(Sys.getenv("ARTICLE_TARGET", "economics")), "nature")) {
  cat("**Keywords**: land tenure; ownership; meta-stochastic frontier; technical efficiency; technology gap; Ghana\n\n")
  cat("**JEL Classification**: Q15, Q12, O13, O55\n")
}
```
