---
title: "Financial Inclusion and Credit Impacts on Crop Production in Ghana"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from financial_inclusion.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: Drafting (v005)](https://img.shields.io/badge/status-drafting-lightgrey)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: Drafting (v005). Consolidated into the canonical study layout on
2026-08-08; the automated narrative system is not yet in place.

### Contributors

- [Mark Appiah-Twumasi](https://scholar.google.com/citations?user=SrQaedsAAAAJ&hl=en)
- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)
- [Aisha Adam](https://scholar.google.com/citations?user=Yf2zSWAAAAAJ&hl=en)
- [Sylvanus Gaku](https://people.extension.wisc.edu/#home/person/66ed8e9bc09f6f031b8fa85e/)
- [Peter Quartey](https://scholar.google.com/citations?user=rEPtTzoAAAAJ&hl=en)

### What this study asks

Financial inclusion — access to savings, credit, digital payments and insurance
— is central to the development strategy of agro-based economies in sub-Saharan
Africa, and recent reports document a marked expansion of (in)formal financial
services in Ghana, including the spread of mobile money. In theory, relaxing
liquidity constraints should let farmers undertake lumpy, yield-enhancing
investments and operate nearer to the production frontier. This study builds a
financial inclusion index for Ghanaian farm households from the Ghana Living
Standards Survey, then applies statistical matching within a meta-stochastic
frontier framework to identify how financial inclusion and credit access affect
technology adoption and technical efficiency in crop production.

<!--
NO ABSTRACT HERE, DELIBERATELY. Once narrative/ is scaffolded it lives in
narrative/sections/00_abstract.Rmd, where every figure is inline R against the
pipeline. A second copy here would be hand-typed and would drift from the paper
-- silently, and on the headline result. Link to it; do not restate it.
-->

The full text, tables and figures are in [`narrative/`](narrative/) — currently
Word drafts (`v005_CreditAgricProdGapGhana.docx`, superseded versions in
`narrative/old/`) pending migration to the automated manuscript system that
`resource_extraction` and `land_tenure` use. **Until that migration lands, the
numbers in the drafts are hand-typed and are not generated from the pipeline.**

**Keywords**: financial inclusion; credit; mobile money; technical efficiency;
meta-stochastic frontier; Ghana

**JEL Classification**: G21, G51, O13, Q12, Q14

### Reproducing

Run from the repository root. There is no `run_article.R` entry point yet — run
the numbered steps directly, in order.

```
000_initialize   scaffolding                       fast
000_INDEX        -> harmonized financial_inclusion_index (Stata)
001_DATA         harmonized releases -> raw data   fast
002_MATCHING     -> estimation_data                EXPENSIVE
003_TREATMENT    -> treatment effects              EXPENSIVE
004_MSF          -> output/estimations/            HPC (job_msf.sbatch)
101_exhibit_*    -> output/figures/                moderate
```

Data preparation upstream of `001` is Stata (`data-raw/scripts/data-prep/glss/03_financial_inclusion.do`),
which harmonizes the raw GLSS files. `scripts/000_INDEX_financial_inclusion_study.do`
builds the financial inclusion index — the treatment variable itself — and
writes it as a harmonized release that `001` reads back.

See [`scripts/README.md`](scripts/README.md) for the naming convention and for
what this study still lacks against `resource_extraction` and `land_tenure`.

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
