---
title: "okwaayeli: Agricultural Productivity Analysis in Ghana"

output:
  github_document:
    toc: false
    toc_depth: 3
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE,
                      echo = TRUE,
                        fig.path = "man/figures/")

if (!requireNamespace("badger", quietly = TRUE)) install.packages("badger")

library(badger)

library(knitr)
library(dplyr)
library(devtools)
devtools::load_all()
devtools::document()
```


<!-- badges: start -->
[![Project Status: Active – The project has reached a stable, usable
state and is being actively
developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/ftsiboe/okwaayeli/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ftsiboe/okwaayeli/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/ftsiboe/okwaayeli/actions/workflows/pkgdown.yaml/badge.svg)](https://ftsiboe.github.io/okwaayeli/)
[![codecov](https://codecov.io/gh/ftsiboe/okwaayeli/graph/badge.svg?token=KIE05NKNU8)](https://codecov.io/gh/ftsiboe/okwaayeli)
![License: GPL v3](https://img.shields.io/badge/License-GPLv3-blue.svg)
![R >= 4.1](https://img.shields.io/badge/R-%3E=4.1-blue)
[![Contributor Covenant](https://img.shields.io/badge/Contributor%20Covenant-2.1-4baaaa.svg)](code_of_conduct.md)
<!-- badges: end -->

*Collaborations with enthusiastic and committed graduate students are welcomed. For more details see the wiki on [Graduate Student Opportunity to Lead Research on Agricultural Productivity in Ghana](https://github.com/ftsiboe/okwaayeli/wiki/Graduate-Student-Opportunity-to-Lead-Research-on-Agricultural-Productivity-in-Ghana).*

## Contents

- [Overview](#-overview)
- [At a glance](#-at-a-glance)
- [Data](#-data)
- [Methods](#-methods)
- [Pipeline](#-pipeline)
- [Published studies](#-published-studies)
- [Work in progress](#-work-in-progress)
- [Scheduled studies](#-scheduled-studies)
- [Installation](#-installation)
- [Quick start](#-quick-start)
- [Contributing](#-contributing)
- [Citation](#-citation)
- [Data and code availability](#-data-and-code-availability)
- [Contact](#-contact)
- [Funding](#-funding)

## 📘 Overview
The primary objective of this project is to determine whether agricultural production shortfalls in Ghana are driven by farmer technical inefficiency, technology gaps, or a combination of both.

It compiles a series of research articles that assess farmer-specific and institutional factors, exploring how they influence technical inefficiency and the adoption of advanced agricultural technologies.

By identifying the root causes of production shortfalls, this project provides empirical evidence crucial for policy discussions aimed at enhancing agricultural productivity in contexts with limited access to new technologies.

## 📌 At a glance

| | |
|---|---|
| **Published outputs** | 9 (7 peer-reviewed articles + 1 data paper + 1 conference paper) |
| **Active manuscript streams** | 8 (disability, education, resource extraction, financial inclusion, land tenure, conflict/peace, income transfer, agricultural services) |
| **Harmonized datasets released** | 11 farmer- and community-level GLSS datasets |
| **GLSS rounds covered** | GLSS1 (1987/88) – GLSS7 (2016/17) |
| **License** | GPL-3 |

## 📊 Data
The project utilizes farm-level data from the Ghana Living Standards Surveys (GLSS), which are population-based surveys conducted periodically, approximately every five years, in Ghana: [GLSS1](https://microdata.statsghana.gov.gh/index.php/catalog/7),  [GLSS2](https://microdata.statsghana.gov.gh/index.php/catalog/4), [GLSS3](https://microdata.statsghana.gov.gh/index.php/catalog/12), [GLSS4](https://microdata.statsghana.gov.gh/index.php/catalog/14), [GLSS5](https://microdata.statsghana.gov.gh/index.php/catalog/5), [GLSS6](https://microdata.statsghana.gov.gh/index.php/catalog/72), and [GLSS7](https://microdata.statsghana.gov.gh/index.php/catalog/97).

The harmonized data products are documented in
[data-raw/releases/harmonized_data/codebook.md](data-raw/releases/harmonized_data/codebook.md)
and distributed through the
[`hh_data` release](https://github.com/ftsiboe/okwaayeli/releases/tag/hh_data).

## 🧪 Methods
- The Meta-Stochastic Frontier (MSF) model, as introduced by [Huang, Huang, & Liu (2014)](https://doi.org/10.1007/s11123-014-0402-2), is employed to capture technological heterogeneity and technical inefficiencies across farmers.
- To minimize self-selection bias and to reduce model dependency, a matching approach is also utilized where applicable ([Abadie and Imbens 2006](https://doi.org/10.1111/j.1468-0262.2006.00655.x); [Abadie and Imbens 2016](https://doi.org/10.3982/ECTA11293)).

For a worked walkthrough of the full pipeline, see the
[methodology vignette](vignettes/msf-and-matching.Rmd).

## 🔗 Pipeline

Every active study follows the same five-stage pipeline. Each stage is a numbered R script whose output feeds the next.

```mermaid
flowchart LR
    A[Harmonized GLSS data<br/>get_household_data()] --> B[001_DATA<br/>sample &amp; covariates]
    B --> C[002_MATCHING<br/>specs &amp; balance]
    C --> D[003_TREATMENT<br/>treatment effects]
    D --> E[004_MSF<br/>frontier estimation]
    E --> F[100_FIGTAB<br/>results.xlsx + figures]
```

## 📄 Published studies
* [Technology and managerial performance of farm operators by age in Ghana](https://doi.org/10.1007/s11123-023-00679-y) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Taking stock of gender gaps in crop production technology adoption and technical efficiency in Ghana](https://doi.org/10.1080/03031853.2022.2150664) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Starchy Staples Production Shortfall in Ghana is influenced by Technical Inefficiency than by Ecological Technology Gaps](https://doi.org/10.1371/journal.pone.0284780) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Production Technology, Efficiency, and Productivity of Cereal Farms](https://doi.org/10.1017/age.2022.16) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Chronic Sources of Low Cocoa Production in Ghana](https://doi.org/10.1017/age.2021.3) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Effect of Fertilizer Subsidy on Household Level Cereal Production in Ghana](https://doi.org/10.1016/j.sciaf.2021.e00916) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Spatiotemporal Evaluation of Dry Beans and Groundnut Production Technology and Inefficiency in Ghana](https://ageconsearch.umn.edu/record/310316/?ln=en&v=pdf) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Vegetable Production Technical Efficiency and Technology Gaps in Ghana](https://ageconsearch.umn.edu/record/301046/?ln=en&v=pdf) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)
* [Nationally Representative Farm/Household Level Dataset on Crop Production in Ghana from 1987-2017](http://dx.doi.org/10.2139/ssrn.4134518) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/releases/tag/legacy_studies)

## 🛠 Work in progress
* [Technology Gaps Drive Inequality in Production for Farmers with Disabilities](https://github.com/ftsiboe/okwaayeli/blob/main/studies/disability/DisabilityAgricProdGapGhana_public_version.pdf) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/disability) — *under review (Agricultural Economics)*
* [Heterogeneous Impacts of Farmer Education on Technology Access and Technical Efficiency in Ghana](https://github.com/ftsiboe/okwaayeli/tree/main/studies/education) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/education) — *revise &amp; resubmit (JAE)*
* [Digging into the Linkages Between Resource Extraction and Farm Performance](https://github.com/ftsiboe/okwaayeli/tree/main/studies/resource_extraction) | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/resource_extraction) — *submitted (Land Use Policy)*
* Societal Peace and Cohesion Improve Farm Technology and Technical Efficiency in Ghana | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/societal_peace_and_cohesion) — *drafting*
* Financial Inclusion and Credit Impacts on Crop Production in Ghana | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/financial_inclusion) — *drafting*
* Land Ownership Appears Neither Necessary nor Sufficient for Superior Agricultural Performance in Ghana | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/land_tenure) — *drafting*
* The Association Between Income Transfers and Crop Production in Ghana | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/) — *concept*
* Agricultural Service Delivery and Smallholder Farm Performance in Ghana | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/ag_services) — *drafting*

## 📅 Scheduled studies
* The link between extension services and Farm Technology Endowment and its Efficient Use — *planned*
* Input-dealer density and farm performance | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/input_dealers) — *code-only, manuscript pending*
* Time poverty and farm output gaps in Ghana | [🗂️⚙️](https://github.com/ftsiboe/okwaayeli/tree/main/studies/time_poverty) — *code-only, manuscript pending*

## ⬇️ Installation

```r
# install.packages("devtools")
devtools::install_github("ftsiboe/okwaayeli")
```

`okwaayeli` requires R ≥ 4.1 and a working C/Fortran toolchain for the
econometric back-ends (`frontier`, `micEcon`, `rgenoud`, `quadprog`,
`optmatch`, `dbarts`, `CBPS`).

## 🚀 Quick start

```r
library(okwaayeli)

# 1. scaffold a project tree under studies/<project_name>/
env <- study_setup(project_name = "tutorial")

# 2. pull a harmonized dataset (cached locally after first call)
farmer_data <- get_household_data("harmonized_crop_farmer_data")

# 3. follow the standard 001 -> 100 pipeline; see the methodology vignette
vignette("msf-and-matching", package = "okwaayeli")
```

## 🤝 Contributing

Please read [CONTRIBUTING.md](CONTRIBUTING.md) for the standard study
pipeline, the "how to add a new study" template, and the coding
conventions. New issues, study proposals and data requests have dedicated
[issue templates](.github/ISSUE_TEMPLATE/). Methodology questions are best
asked under [Discussions](https://github.com/ftsiboe/okwaayeli/discussions).

This project follows the [Contributor Covenant](code_of_conduct.md).
Security disclosures: see [SECURITY.md](SECURITY.md).

## 📚 Citation

To cite the package itself:

```r
citation("okwaayeli")
```

For the harmonized dataset underlying every study:

> Tsiboe, F. (2022). *Nationally Representative Farm/Household Level
> Dataset on Crop Production in Ghana from 1987–2017.* SSRN Electronic
> Journal. <https://doi.org/10.2139/ssrn.4134518>

Please also cite the specific paper(s) whose data or method you use; the
full list is in [Published studies](#-published-studies).

## 🔓 Data and code availability
The datasets and code used in this project are openly available to facilitate replication and further research. They may also be used freely for related studies. To help track the usage and impact of these resources, please inform us if you intend to use them beyond direct replication.

## 📬 Contact
Constructive feedback is highly appreciated, and collaborations using the available codes and data are actively encouraged. Please reach out by sending emails to Francis Tsiboe ([ftsiboe@hotmail.com](mailto:ftsiboe@hotmail.com)).

## 💰 Funding
This project did not receive any external funding.

See the [LICENSE](LICENSE) file in the repository's root for details.

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
