---
title: "Meta-stochastic frontiers and statistical matching: the okwaayeli pipeline"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meta-stochastic frontiers and statistical matching: the okwaayeli pipeline}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  eval     = FALSE
)
```

## Why this pipeline exists

Sub-Saharan crop yields are typically 30–60% below global benchmarks. The
shortfall has two distinct sources:

- **Technical inefficiency** — farmers operating *below* the best practice
  that is already available in their district.
- **Technology gap** — entire regions (or sub-populations) lacking the
  inputs, varieties, or know-how that define the global / national
  production frontier.

The two sources call for different policy responses. Closing inefficiency
mainly needs extension and managerial support; closing technology gaps
needs investment in inputs, varieties, and infrastructure. To decide where
scarce policy attention should go, you need to *decompose* the observed
output shortfall.

`okwaayeli` implements this decomposition through a meta-stochastic-frontier
(MSF) framework (Huang, Huang, and Liu, 2014) coupled with statistical
matching (Abadie & Imbens, 2006, 2016) so that group comparisons (e.g.
farmers with vs without disabilities) are not contaminated by selection on
observables.

## The pipeline at a glance

```
┌──────────────┐   ┌─────────────────┐   ┌──────────────────┐   ┌────────────┐   ┌────────────┐
│  Harmonized  │   │  001_DATA       │   │  002_MATCHING    │   │  003_TE    │   │  004_MSF   │
│  GLSS data   ├──▶│  (sample        ├──▶│  (write specs,   ├──▶│  (treat-   ├──▶│  (frontier │
│  (release)   │   │   restriction,  │   │   match, balance)│   │   ment     │   │   estim'n) │
└──────────────┘   │   covariates)   │   └──────────────────┘   │   effects) │   └─────┬──────┘
                   └─────────────────┘                          └────────────┘         │
                                                                                       ▼
                                                                       ┌────────────────────────────┐
                                                                       │  100_FIGTAB                │
                                                                       │  (figures, tables,         │
                                                                       │   *_results.xlsx)          │
                                                                       └────────────────────────────┘
```

Each study under `replications/<study>/` follows this same skeleton.

## Step 0 — set up

```{r}
library(okwaayeli)

study_environment <- study_setup(project_name = "tutorial")
```

`study_setup()` creates the standard folder tree under
`replications/<project_name>/output/` and returns the resolved paths plus
a reproducible seed (default `11122025`, set via `okwaayeli_control()`).

## Step 1 — pull harmonized data (001_DATA)

```{r}
farmer_data     <- get_household_data("harmonized_crop_farmer_data")
disability_data <- get_household_data("harmonized_disability_data")

study_data <- merge(
  farmer_data, disability_data,
  by = c("Surveyx", "EaId", "HhId", "Mid")
)
```

`get_household_data()` downloads the requested `.dta` file once from the
`hh_data` GitHub release via `piggyback`, caches it under
`tools::R_user_dir("okwaayeli", "cache")`, and returns a data.frame.

Typical 001 work after the merge: restrict the survey waves, build the
treatment indicator (e.g. `disabled <- as.integer(disabled_self == 1 | …)`),
build the covariate list.

## Step 2 — matching (002_MATCHING)

```{r}
formulas <- write_match_formulas(
  treatment = "disabled",
  covariates = c("Female", "Age", "YerEdu", "HouseholdSize", "Region")
)

matched <- match_sample_specifications(
  data     = study_data,
  formulas = formulas,
  weight   = "Weight"
)

balance <- covariate_balance(matched)
```

The matching engine evaluates several specifications (nearest-neighbour,
CBPS, propensity-score, BART, etc.) in parallel and reports standardised
mean differences so you can pick the best-balanced one. Output is written
to `output/matching/`.

## Step 3 — treatment effects (003_TREATMENT)

```{r}
draws <- draw_matched_samples(matched, n_draws = 100)

te_summary <- treatment_effect_calculation(
  draws  = draws,
  data   = study_data,
  outcomes = c("HrvstKg", "FertKg", "SeedKg")
)

treatment_effect_summary(te_summary)
```

Outputs land in `output/treatment_effects/` and feed both the MSF
estimation and the final exhibits.

## Step 4 — meta-stochastic frontier (004_MSF)

```{r}
msf_specs <- list(
  output_variable  = "HrvstKg",
  input_variables  = c("Area", "LabHr", "FertKg", "SeedKg", "PstcdHr"),
  inefficiency_covariates = list(
    Svarlist = c("YerEdu", "Female", "Region"),
    Fvarlist = c("Surveyx")
  ),
  risk_covariates  = list(Svarlist = c("Region")),
  weight_variable  = "Weight"
)

est <- draw_msf_estimations(
  draw                 = 0,
  drawlist             = draws,
  data                 = study_data,
  output_variable      = msf_specs$output_variable,
  input_variables      = msf_specs$input_variables,
  inefficiency_covariates = msf_specs$inefficiency_covariates,
  risk_covariates      = msf_specs$risk_covariates,
  weight_variable      = msf_specs$weight_variable
)

summary_est <- draw_msf_summary(est, technology_legend = c("With", "Without"))
```

The MSF block returns input elasticities, returns to scale, technical
efficiency (TE), technology-gap ratios (TGR), and the meta-frontier
performance metric (TE × TGR). Outputs land in `output/estimations/`.

For real studies you typically run draws 1..N in parallel via the SLURM
`job_msf.sbatch` script that ships in each replication folder, then
aggregate.

## Step 5 — figures and tables (100_FIGTAB)

The final stage writes the canonical `*_results.xlsx` exhibit plus the
publication-ready figures into `output/figure/` and `output/figure_data/`.
The shell of this script is study-specific because exhibits vary, but the
inputs are always the per-draw estimation `.rds` files written in step 4.

## What to read next

- `?study_setup` and `?okwaayeli_control` for the project scaffolding.
- `?get_household_data` for the data download mechanics.
- `?msf_workhorse` and `?sf_workhorse` for the core estimators.
- The `disability` study under `replications/disability/` for a fully
  worked example.

## References

- Huang, C. J., Huang, T. H., & Liu, N. H. (2014). A new approach to
  estimating the metafrontier production function based on a stochastic
  frontier framework. *Journal of Productivity Analysis*, 42(3), 241–254.
  <https://doi.org/10.1007/s11123-014-0402-2>
- Abadie, A., & Imbens, G. W. (2006). Large sample properties of matching
  estimators for average treatment effects. *Econometrica*, 74(1), 235–267.
  <https://doi.org/10.1111/j.1468-0262.2006.00655.x>
- Abadie, A., & Imbens, G. W. (2016). Matching on the estimated propensity
  score. *Econometrica*, 84(2), 781–807.
  <https://doi.org/10.3982/ECTA11293>
