---
title: "Extractive Land Uses and Farm Performance: Evidence of Technology Gaps and Adaptive Efficiency"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from resource_extraction.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: Submitted (Land Use Policy)](https://img.shields.io/badge/status-submitted-yellow)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: Submitted to Land Use Policy (cover letter dated 17 May 2026).

### Contributors

- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)
- [Mark Appiah-Twumasi](https://scholar.google.com/citations?user=SrQaedsAAAAJ&hl=en)
- [Antoinette Tsiboe-Darko](https://csps.ug.edu.gh/antoinette-adjoa-tsiboe-darko-phd)
- [Sylvanus Gaku](https://people.extension.wisc.edu/#home/person/66ed8e9bc09f6f031b8fa85e/)

### Abstract

Ensuring sustainable productivity in Ghana's agriculture is essential for national food security and poverty reduction, yet rapid expansion of resource extraction poses new challenges. This paper examines how proximity to extraction sites influences farmers' use of production technologies and technical efficiency. Using the latest four waves of the Ghana Living Standards Survey together with community-level extraction indicators, statistical matching, and a meta-stochastic frontier framework, we distinguish technology gaps from technical efficiency among farmers exposed to gold and bauxite mining, sand winning, and stone quarrying. We find that farmers in extraction communities generally operate with weaker production technologies but often use the technologies they do have more efficiently, with important differences across extraction types. The results suggest that land-use policy should avoid treating extractive activities as a single policy problem and should instead target the specific combination of technology access and managerial constraints each type of extraction creates.

**Keywords**: resource extraction; mining; technology gap; technical efficiency; meta-stochastic frontier; Ghana

**JEL Classification**: Q15, Q32, Q33, O13, Q12

### Project layout

This study follows the standard research-project framework:

```
resource_extraction/
  scripts/            analysis pipeline + article layer
    000_initialize.R          bootstrap: source helpers, build output/ tree, seed
    001_DATA_…_study.R        assemble the analysis dataset
    002_MATCHING_…_study.R    proximity-based matching
    003_TREATMENT_…_study.R   treatment-effect construction
    004_MSF_…_study.R         meta-stochastic frontier estimation
    100_exhibits.R (.do)      build figures/tables into output/exhibits/
    300_article_helpers.R     paths, fmt_* helpers, assert_present()
    301_article_objects.R     assemble narrative/article_objects.json
    302_render_article.R      knit the master narrative to .docx/.html
    run_article.R             source 300 -> 301 -> 302
    job_msf.sbatch            HPC batch job for 004_MSF
  narrative/          the article, split into reproducible sections
    resource-extraction.Rmd   master Rmd (stitches sections/*.Rmd)
    sections/                 00_abstract … 98_tables_and_figures
    references.bib            bibliography seed
    reference.docx            Word style reference
    review_comments.Rmd       reviewer-comment tracker
    writeup/ submissions/ literature/   private drafts, journal packages, source PDFs
  data/               study inputs (gitignored)
  output/             estimations/ matching/ exhibits/ summary/ releases/ (gitignored)
```

### Analysis pipeline

Run from the repository root, in order:

1. `scripts/001_DATA_resource_extraction_study.R` — build the analysis dataset.
2. `scripts/002_MATCHING_resource_extraction_study.R` — match extraction vs. non-extraction farmers.
3. `scripts/003_TREATMENT_resource_extraction_study.R` — construct treatment effects.
4. `scripts/004_MSF_resource_extraction_study.R` — estimate the meta-stochastic frontier (or submit `scripts/job_msf.sbatch`).
5. `scripts/100_exhibits.R` — build figures and tables into `output/exhibits/`.
6. `scripts/run_article.R` — assemble article objects and knit the manuscript.

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
