---
title: "Landownership Appears Neither Necessary nor Sufficient for Superior Agricultural Performance in Ghana"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from land_tenure.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: In revision (v002)](https://img.shields.io/badge/status-in%20revision-blue)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: v002 in preparation. v001 was produced 20 May 2026; a GAAE 2025
conference paper is under `submissions/conferences/`.

### Contributors

- [Gilbert Addae](https://github.com/Gilbertaddae)
- [Jacob Asravor](https://scholar.google.com/citations?user=_zUi3FsAAAAJ&hl=en)
- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)

### What this study asks

Secure land tenure is widely assumed to stimulate investment and raise
agricultural productivity, yet rigorous evidence from sub-Saharan Africa
remains inconclusive. Using five rounds of the Ghana Living Standards Survey
(GLSS 3-7), this study links land-ownership status to production outcomes,
applying statistical matching within a meta-stochastic frontier framework to
isolate ownership effects on input elasticities, returns to scale, technology
gaps, and technical efficiency. It then disaggregates by tenure form —
documentation, transfer rights, mode of acquisition, and sharecropping
intensity.

<!--
NO ABSTRACT HERE, DELIBERATELY.

This file used to carry a hand-copied abstract with hard-coded figures. It
drifted: it claimed six GLSS rounds (there are five), froze N at 35,185 and the
returns-to-scale gap at 0.0154, listed JEL K11 where the paper lists O55, and
reported "no statistically significant difference" where the manuscript reports
"statistically significant but economically small differences" -- the opposite
conclusion, on the study's headline result.

The abstract lives in narrative/sections/00_abstract.Rmd, where its numbers are
inline R against the pipeline and cannot go stale. A README is not the place to
keep a second copy of a result. Link to it; do not restate it.
-->

The abstract, full text, tables and figures are in
[`narrative/`](narrative/) — see `narrative/land-tenure.Rmd` and the rendered
`land-tenure.docx`. Every number in the manuscript is generated from the
pipeline; none is typed by hand.

**Keywords**: land tenure; ownership; meta-stochastic frontier; technical
efficiency; technology gap; Ghana

### Reproducing

Run from the repository root. `scripts/run_article.R` is the single entry
point; set a stage to `TRUE` to run it.

```
000_initialize   scaffolding                       fast
001_DATA         harmonized releases -> raw data   fast
002_MATCHING     -> estimation_data                EXPENSIVE
003_TREATMENT    -> treatment effects              EXPENSIVE
004_MSF          -> output/estimations/            HPC (job_msf.sbatch)
100_exhibit_*    -> data/descriptive_exhibits.rds  ~5-10 min
101_exhibit_*    -> output/figures/                moderate
102_exhibit_*    -> output/tables/*.xlsx           fast
301/302          -> narrative/land-tenure.docx     fast
```

Data preparation upstream of `001` is Stata (`data-raw/okwaayeli_DATA.do`),
which harmonizes the raw GLSS files. Everything from `001` onward is R: all
tables and figures are built from the pipeline, with one deliberate exception —
Table S0 documents how the tenure indicators were constructed from each round's
questionnaire, and is transcribed rather than computed.

See [`scripts/README.md`](scripts/README.md) for the naming convention and
[`narrative/diagnostics/`](narrative/diagnostics/) for the variable-construction
audit trail.

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
