---
title: "Time Poverty and Farm Output Gaps in Ghana"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from time_poverty_study.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: Code-only (manuscript pending)](https://img.shields.io/badge/status-code--only-lightgrey)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: Code-only at this stage. Replication scripts (001-005 plus 100_FIGTAB) are in place; the working paper is pending.

### Contributors

- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)

### Abstract

Unpaid domestic and care work absorbs a disproportionate share of women's time in rural Ghana and may translate into measurable losses in farm output through reduced labour effort, reduced adoption of labour-saving technologies, and reduced managerial attention. Using the GLSS7 time-use harmonization (`harmonized_time_poverty_data.dta` plus the paid/unpaid companion datasets), statistical matching, and a meta-stochastic frontier framework, this study estimates how time-poverty affects technology adoption and technical efficiency among Ghanaian crop farmers, with a particular focus on the gender dimension. *Full abstract to follow once v000 lands.*

**Keywords**: time poverty; gender; unpaid work; technical efficiency; meta-stochastic frontier; Ghana

**JEL Classification**: J16, J22, O13, Q12

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
