---
title: "Production Gaps Among Disabled Farmers Are Associated with Limited Technology"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from disability_study.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: Under review (Agricultural Economics, R2)](https://img.shields.io/badge/status-under%20review-blue)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: Under review at Agricultural Economics (round 2). Working paper:
[DisabilityAgricProdGapGhana_public_version.pdf](https://github.com/ftsiboe/okwaayeli/blob/main/replications/disability/DisabilityAgricProdGapGhana_public_version.pdf).

### Contributors

- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)
- [Edward Martey](https://scholar.google.com/citations?user=SMoBI3wAAAAJ&hl=en)
- [Jacob Asravor](https://scholar.google.com/citations?user=_zUi3FsAAAAJ&hl=en)

### Abstract

In sub-Saharan Africa, agriculture is vital; however, the systematic exclusion of persons with disabilities from farming hinders sustainable development and social equity. While several studies have analyzed productivity differentials across various dimensions of social exclusion, empirical evidence on technology adoption, farming efficiency, and productivity gaps among farmers with disabilities remains limited. This study addresses this gap by evaluating disparities in crop production between farmers with and without disabilities in Ghana, with particular focus on differences in technology access and production efficiency. Utilizing meta-stochastic frontier analysis coupled with statistical matching techniques on nationally representative farm-level data from the 2012/13 and 2016/17 agricultural seasons, we find negligible differences in technical efficiency between the two groups (less than 0.5%). However, the productive capacity of the technology sets employed by farmers with disabilities is approximately 11 percentage points lower than that of their counterparts without disability, resulting in a 9.5 percentage points shortfall in crop production. This production shortfall is primarily driven by reduced per-hectare use of critical inputs such as planting materials, labor, fertilizer, and agro-chemicals. These findings highlight a significant disability-induced technology access gap which reflects broader structural inequalities. Addressing these disparities is essential for promoting equitable agricultural development while harnessing the full potential of all farmers in Ghana.

**Keywords**: disability parity; agricultural technology; technical efficiency; Ghana; production gap

**JEL Classification**: J14, Q12, O13, O33, I38

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
