---
title: "Digging into the Linkages Between Resource Extraction and Farm Performance"
author:
  - Francis Tsiboe
  - Mark Appiah-Twumasi
  - Antoinette Tsiboe-Darko
  - Sylvanus Gaku
output:
  # officedown (not word_document) so exhibits can sit in landscape sections;
  # word_document has no mechanism for page orientation.
  #
  # fig_width/fig_height: Word output defaults to 5x4in, which knitr stamped onto
  # every figure -- ignoring out.width AND the source aspect ratio (an 11x7 plot
  # came out 5x4, i.e. squashed). The exhibits are now raw markdown images with
  # pandoc {width=} attributes, so these defaults no longer bite, but they are set
  # to the portrait text width here so anything added later starts sane.
  officedown::rdocx_document:
    reference_docx: reference.docx
    fig_width: 6.3
    fig_height: 4.5
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    css: css/tables.css
bibliography: references.bib
# Citation style — switch by changing the `csl:` line below (files live in csl/):
#   Food Policy (Elsevier Harvard, author-date):  csl/elsevier-harvard.csl
#   IEEE (numbered):                               csl/ieee.csl
csl: "`r Sys.getenv('ARTICLE_CSL', 'csl/elsevier-harvard.csl')`"
link-citations: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
if (file.exists("article_objects.json")) {
  objs <- jsonlite::fromJSON("article_objects.json")
}
# Table builders (flextable); paths are repo-root relative.
if (file.exists("../scripts/305_tables.R")) source("../scripts/305_tables.R")
```

Francis Tsiboe -- Agricultural Risk Policy Center, North Dakota State University, Fargo, ND 58102

Mark Appiah-Twumasi[^corresp] -- Department of Agricultural and Food Economics, FoAFCS, University for Development Studies, Tamale, Ghana

Antoinette Tsiboe-Darko -- Centre for Social Policy Studies (CSPS), University of Ghana, Legon

Sylvanus Gaku -- Division of Extension, University of Wisconsin-Madison, Madison, WI 53705

[^corresp]: Corresponding author. Email: mtwumasi@uds.edu.gh

```{r child = "sections/00_abstract.Rmd"}
```

```{r child = "sections/01_introduction.Rmd"}
```

```{r child = "sections/02_data.Rmd"}
```

```{r child = "sections/03_state_of_extraction.Rmd"}
```

```{r child = "sections/04_methods.Rmd"}
```

```{r child = "sections/05_results.Rmd"}
```

```{r child = "sections/06_discussion.Rmd"}
```

```{r child = "sections/07_conclusion.Rmd"}
```

```{r child = "sections/96_declarations.Rmd"}
```

```{r child = "sections/98_tables_and_figures.Rmd"}
```

```{r child = "sections/97_references.Rmd"}
```

```{r child = "sections/99_appendix.Rmd"}
```
