---
title: "Input-Dealer Density and Farm Performance in Ghana"

output:
  github_document:
    toc: false
    toc_depth: 3

knit: >
  (function(inputFile, encoding) {
    rmarkdown::render(
      inputFile,
      encoding      = encoding,
      output_format = "github_document",
      output_file   = paste0(getwd(), "/README.md")
    )
  })
---

<!-- README.md is generated from input_dealers.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", out.width = "100%")
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE, echo = TRUE)
```

![Status: Code-only (manuscript pending)](https://img.shields.io/badge/status-code--only-lightgrey)

For an overview of the broader project context, please refer to the main
[okwaayeli README](https://github.com/ftsiboe/okwaayeli/blob/main/README.md)
in the repository root.

See the
[LICENSE](https://github.com/ftsiboe/okwaayeli/blob/main/LICENSE)
file in the repository root for details.

**Status**: Code-only at this stage. Replication scripts (001-004 plus a threshold-search routine) are in place; the working paper is pending.

### Contributors

- [Francis Tsiboe](https://scholar.google.com/citations?user=ox2t_YIAAAAJ&hl=en)

### Abstract

Access to fertilizer, seed, and agrochemical retailers is a binding constraint on technology adoption for many smallholder farmers in Ghana. This study links the geocoded agro-dealer sale points dataset (`ghana_agro_dealer_sale_points.rds`) to harmonized GLSS farmer data and uses a threshold-search routine plus a meta-stochastic frontier framework to test whether proximity to input dealers raises farm-level technology adoption and technical efficiency. *Full abstract to follow once v000 lands.*

**Keywords**: agro-dealers; input access; technical efficiency; meta-stochastic frontier; Ghana

**JEL Classification**: Q12, Q13, O13, L81

---

*Maintained by [ftsiboe](https://github.com/ftsiboe)*
